---
title: "FM-PreRegistration"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---
# load in tidyverse
```{r}
library(tidyverse)
```

# read in data from sona and prolific
```{r}
sona_data = read.csv("../data/savic-rhyming-sona.csv") %>%
  select(-sona_id) %>% mutate(data_source = "sona")

# repeat for prolific

prolific_data = read.csv ("../data/savic-rhyming-prolific.csv") %>%
  select(-PROLIFIC_PID) %>% mutate(data_source = "prolific")

```
# Making sure sona data and prolific data have the same amount of columns
```{r}
ncol(sona_data)
ncol(prolific_data)
prolific_data <- prolific_data %>% select(names(sona_data))
```

# bind
```{r}
##prolific_data <- prolific_data %>% select(names(sona_data))
final_data = rbind(sona_data, prolific_data) %>%
 mutate(rt = as.numeric(rt),
        relatedness = as.factor(relatedness),
        condition = as.factor(condition))%>%
  mutate(condition = fct_recode(condition, 
                                "non-rhyming" = "1",
                               "rhyming" = "2"))
```
#INSPECT DATA
 
```{r}
#number of rows and columns 
nrow(final_data)
ncol(final_data)
```
#Number of columns = 49
#Number of rows = 44306


```{r}
#unique participants and trials

final_data%>% 
group_by(ID)%>% 
count()

final_data %>% 
  filter(typeoftrial == "sentence") %>%
  group_by(ID) %>% 
  count()

final_data %>% 
  filter(typeoftrial == "attention") %>%
  group_by(ID) %>% 
  count()

final_data %>% 
  filter(typeoftrial == "target") %>%
  group_by(ID) %>% 
  count()

```
#There are 136 unique subjects but only have complete data for 126 participants.
#Subjects completed different numbers of trials because _________
# 60 sentence trials were completed all participants except 232084287, 467885230, 727137633, 869755378.		
# 9 attention check trials were completed by all participants except 232084287, 467885230, 727137633, 869755378, 
# All participants completeted 55 trials; however, these 4 subject IDs did not appear in the data because they had no target data.


# Levels of variables
```{r}
levels(final_data$condition)
levels(final_data$relatedness)
```
#Our Condition and Relatedness columns indicate our independent variables.
#Condition has 2 levels: non-rhyming and rhyming
#Relatedness has 2 levels that we are evaluating (related and unrelated). In the data it has "novel" and "" too. No target trial has "". 

#Our dependent variable columns are reaction time and correct for target trials. For our attention check and association task, we are looking at the response column. 

#BASIC DESCRIPTIVES
```{r}
demographics= final_data %>%
  filter(typeoftrial == "demographics") %>%
  select("ID","gender","age","education","race","hispanic","dominant_hand","alert_time","english")%>%
  mutate(across(c("age","education"), ~ replace_na(., NA)))%>%
 mutate(across(c("gender","race","hispanic","dominant_hand","alert_time","english"),
                 ~ replace_na(., "NOT_FOUND"))) %>%
mutate(across(c("age","gender","education","race","hispanic", "dominant_hand", "alert_time", "english"),
              ~ifelse(. == "", "blank",.)))

subject_age= demographics%>%
  summarise(mean_age= mean(age, na.rm = TRUE),
            sd_age=sd(age, na.rm = TRUE))

gender_distribution= demographics%>%
  filter(gender != "blank") %>%
  count(gender)

race_distribution= demographics%>%
  summarise(race != "blank")%>%
  count(race)

education_distribution= demographics%>%
  summarise(education != "blank")%>%
  count(education)
```
# Basic Descriptives cont'd
#What is the average accuracy in your experiment? What is the standard deviation of the accuracy?
```{r}
trial_accuracy= final_data%>%
filter(typeoftrial == "target") %>%
  summarise(mean_accuracy = mean(correct), 
            sd_accuracy = sd(correct))

## summarize participant accuracy 
subject_attention_accuracy = attention_trials  %>% 
  group_by(ID)  %>% 
  summarise(mean_accuracy = mean(correct))

# find IDs with less than 75% accuracy
low_acc_IDs = subject_attention_accuracy  %>% 
  filter(mean_accuracy < 0.75)  %>% 
  pull(ID)
```
#still trying to get the histogram to work
```{r}
basic_data = final_data %>% filter(typeoftrial == "target") %>% 
  select (ID, rt, condition, relatedness, prime, response, type, correct, block_number, target, correct_key) %>% 
ggplot() +
  geom_histogram(mapping = aes (x = as.numeric(rt)))

```

#INFERENTIAL STATISTICS

#Our primary research question is whether exposure to language containing rhyming words can help to foster semantic links between novel and familiar words. If rhyming facilitates semantic integration, participants should identify familiar nouns faster when the prime is related to and  rhymes with the familiar noun.

#In our pre-registration, we said that we would exclude trials that accuracy on attention check questions less than 0.75 will result in data being excluded from previous analyses. On the priming task, incorrect identification of the familiar noun or reaction times that are extremely short (less than 200 ms) or extremely long (greater than 1,500 ms) will result in removal of data. In addition, incomplete experiments will be removed from analyses. Finally, English fluency is necessary, so data will be removed if a participant’s first language is not English or they did not learn before the age of 4. 

```{r}
#editing all attention trials to adjust for typos

attention_trials= final_data%>% 
  filter(typeoftrial == "attention") %>% 
  select(ID, response, novel1, novel2, correct) %>% 
    rowwise() %>%
  mutate(response = ifelse(is.na(response), "blank", response)) %>%
  mutate(across(c(novel1, novel2), ~ replace_na(., "NOT_FOUND"))) %>% 
  mutate(edit_novel1 = adist(novel1, response),
         edit_novel2 = adist(novel2, response)) %>% 
  mutate(revised_correct = ifelse(edit_novel1 <= 2 |
                                    edit_novel2 <= 2,
                                    1, 0),
          mismatch = ifelse(correct == revised_correct, 0, 1)) %>%
  ungroup()
```



```{r}
priming_data = final_data %>% filter(typeoftrial == "target") %>% 
  select (ID, rt, condition, relatedness, prime, response, type, correct, block_number, target, correct_key) %>% 
  filter(!is.na(rt), rt> 200, rt< 1500, correct == "true", block_number == 1) %>% 
  filter(relatedness %in% c("related", "unrelated")) %>%
  filter(!ID %in% low_acc_IDs)

```

```{r}
final_data %>%
  group_by(condition, relatedness) %>%
  summarise(mean_rt = mean(rt)) %>%
  ggplot() +
  geom_col(mapping= aes(x= condition, y= mean_rt, 
                        group= relatedness, fill= relatedness),
           position = "dodge") +
  theme_bw() +
  scale_fill_manual(values = c("firebrick2", "lightblue"))
```

```{r}
library(lmerTest)

rt_model = lmer(data = final_data, rt~ relatedness*condition + (1|ID))
```

```{r}
car::Anova(rt_model)
nobs(rt_model)
# load packages
check_model(rt_model)
```





